Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming

### 2017-04-25 docs/paperin_review.tex
tweak plotting functions to accomodate modeling and fitting changes

\usepackage{tikz}
\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}

@@

$c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\

@@

$c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\

@@

$e$ \> Oswald efficiency factor \\

@@

$m_{\text{fac-wing}}$ \> margin factor\\
$m_{\text{fac-wing}}$ \> wing weight margin factor \\

@@

$\mathcal{S}$ \> shear forces \\ % [N] \\

@@

@@

This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two archictures for a given mission.  Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
The optimization is solved multiple times to observe how aircraft size and weight are affected by different design parameters and requirements.

@@

Using this methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
Similar trade offs are calculated for other requirements and performance metrics to determine under what conditions each platform is best suited to meet a set of requirements.

@@

Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
Therefore, the station keeping requirement is defined as a percentage which is the percentile wind speed. 
The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,

@@

For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement. 
Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.B, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.  
Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June will be used for the aircraft sizing and performance analysis presented in this paper.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Winds peak at summer and winter solstice in Southern and Northern Hemispheres respectively.\cite{wind}}}
 \label{f:windvsmonth}
\end{figure}

@@

Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m. 
Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m depending on the engine.  
Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.

@@

Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months. This is further discussed in Section IV.B.

@@

Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
Common characteristics between the two models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
The solar-electric powered aircraft holds the batteries in the wings.  
The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.

@@

These assumptions can be changed to evaulate different configuration.  Changes in configuration assumptions may require modification of models or equations.

@@

Because wind speeds also vary with latitude, several constraints are needed to optimize altitude. 
Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg et. al\cite{fitting}.
Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
Altitude can be post-calculated from the optimized air density.

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
% The constraint on December wind speeds and air density at the 30 degree latitude is 
% 
% \begin{align}
%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
%     \label{e:windfit35}
%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
% \end{align}
% 
% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed. The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}. 
% 
% \begin{figure}[H]
% 	\begin{center}
% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
% 	\label{f:windfitl35}
% 	\end{center}
% \end{figure}

@@

\begin{align}
    \label{e:solarreq}
    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
    \label{e:solarbatt}
    E_{\text{batt}} &\geq \frac{P_{\text{oper}}t_{\text{night}}}{\eta_{\text{discharge}}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
\end{align}

@@

where $\eta_{\text{motor}} = 0.95$. Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude. 
$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix B to generate the fitted equations. 
% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
% 
% \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
%  \end{subfigmatrix}
%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
%  \label{f:energyapprox}
% \end{figure}

@@

@@

This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
$W_{\text{ave}}$ is assumed to be the geometric mean, defined as

@@

    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix B for a discussion of the airfoil choice). 
    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.

    \begin{align}
        \label{e:aerodragprof}
        c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
                       &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
        Re &= \frac{\rho V S/b}{\mu}
    \end{align}

@@

\item[Standard Wing Bending Case]

@@

\item[Gust Loading Case]

@@

Because the gust velocity is assumed to be vertical to the flight path, the local angle of attack is approximated by

@@

\item[Discretized Beam Model]

@@

\item[Cap Spar for Bending Loads]

A cap spar is considered for the solar-electric and gas powered aircraft.  Both aircraft optimization models share the same governing equations.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.

@@

\subsubsection{Wing Skin Weight}

@@

@@

The gas and solar-electric powered aircraft optimization models were solved by minimizing take of weight for the requirements listed in Table~\ref{t:mreqs}. 
Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take off weight of 84.0 lbs.  
The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take off weight of 320.5 lbs.
According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.

\begin{longtable}{llll}
\caption{Fixed Optimization Parameters} \\
\toprule
\toprule
\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
\midrule
$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
$\tau$                       & 0.5  & $\tau$                       & 0.115 \\
$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
$\dot{h}$ [ft/min]           & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
                             &     & $\eta_{\text{charge}}$          & 0.98 \\
                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
\bottomrule
\label{t:gassolarparams}
 \end{longtable}

@@

Important trade offs between the gas powered and solar-electric powered archictures are highlighted and quantified by changing the latitude and endurance endurance requirements.   
Both optimization models were solved by minimizing the max take off weight across different bands of latitudes. 
Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
The gas powered model was solved x times and the solar-electric powered model was solved x times to produce Figure~\ref{f:latvsmtowtrade}.
The gas powered optimization model contains 606 free variables and the solar-electric powered optmization model contains between 165 and 585 free variables depending on latitude band requirement.

@@

While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
Figure~\ref{f:spanvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
	\label{f:spanvsendurance}
	\end{center}
\end{figure}

@@

The previous results are dependent on the assumed input values and parameters.  
Changing the input values can help show where the designs becomes infeasible. 
Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude.

@@

By altering the aircraft structural model it can be observed how air density trades for wing weight. 
It might be assumed that because the wind speeds are lowest at 67,000 ft at 29 degrees latitude, that the aircraft will always fly at 67,000 ft.  
If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight

@@

\section*{Appendix A}

\subsection{Wind Speed Constraints}

The each wind speed constraint follows the following form

\begin{equation}
    \label{e:windspeedgen}
    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{\alpha} \geq c_1 \rho^{e_{1,1}}p_{\text{wind}}^{e_{1,2}} + c_2 \rho^{e_{2,1}}p_{\text{wind}}^{e_{2,2}} + c_3 \rho^{e_{3,1}}p_{\text{wind}}^{e_{3,2}} + c_4 \rho^{e_{4,1}}p_{\text{wind}}^{e_{4,2}}
\end{equation}

Table~\ref{t:windvals} lists the values of the coefficients and exponents of Equation~\eqref{e:windspeedgen} for latitudes 20-60. Each generated constraint has a RMS error of less than 5\% valid for altitude ranges between 48,000 and 80,000 ft.

\tiny
\begin{longtable}{lccccccccccccc} \\
\label{t:windvals} \\
    \caption{Wind Speed Function Coefficients and Exponents} \\
    \toprule
    \toprule
    Latitude & $c_1$ & $e_{1,1}$ & $e_{1,2}$ & $c_2$ & $e_{2,1}$ & $e_{2,2}$ & $c_3$ & $e_{3,1}$ & $e_{3,2}$ & $c_4$ & $e_{4,1}$ & $e_{4,2}$ & $\alpha$ \\
    \midrule
20 & 9.77e-15 & -8.45 & 53.1 & 5.26e+04 & 9.98 & 6.26 & 1.97e+03 & 7.66 & 39.3 & 3.9e-14 & -7.26 & 10.6 & 5.53\\
21 & 1.01e-13 & -7.88 & 47.9 & 6.66e-13 & -6.62 & 9.18 & 8.09e+04 & 9.61 & 5.48 & 1.36e+03 & 7.04 & 35.8 & 5.03\\
22 & 1.12e+05 & 9.65 & 5.29 & 1.53e-13 & -7.75 & 47.6 & 7.58e-13 & -6.64 & 8.9 & 2.03e+03 & 7.11 & 35 & 5.06\\
23 & 2.68e+03 & 7.11 & 34 & 1.07e-12 & -6.61 & 8.67 & 1.35e+05 & 9.6 & 5.09 & 4.2e-13 & -7.46 & 48.6 & 5.08\\
24 & 4.81e+03 & 7.4 & 33.6 & 6.8e-13 & -7.25 & 48.3 & 1.95e+05 & 9.81 & 5.08 & 4.84e-13 & -6.87 & 8.91 & 5.31\\
25 & 2.14e-12 & -6.86 & 48 & 6.28e+03 & 7.5 & 33.3 & 2.02e+05 & 9.8 & 5.12 & 3.1e-13 & -7.06 & 9.17 & 5.48\\
26 & 7.48e+03 & 7.51 & 34.7 & 2.95e+05 & 9.9 & 5.26 & 3.13e-13 & -7.14 & 9.52 & 7.03e-12 & -6.51 & 51.4 & 5.59\\
27 & 3.62e+05 & 10 & 5.44 & 7.09e-12 & -6.54 & 51.1 & 1.29e-13 & -7.44 & 9.71 & 8.93e+03 & 7.57 & 36.3 & 5.82\\
28 & 7.9e-14 & -7.65 & 9.68 & 8.89e+03 & 7.5 & 36.9 & 3.28e+05 & 9.95 & 5.58 & 6.85e-12 & -6.64 & 49.3 & 5.96\\
29 & 6.97e-14 & -7.79 & 9.85 & 2.3e+05 & 9.71 & 5.72 & 7.67e+03 & 7.34 & 36.4 & 1.15e-11 & -6.58 & 50.3 & 6.02\\
30 & 4.2e-14 & -8.05 & 10.2 & 1.29e+05 & 9.38 & 6 & 5.8e+03 & 7.15 & 36.4 & 1.44e-11 & -6.62 & 52.2 & 6.12\\
31 & 2.17e-12 & -7.2 & 57.6 & 9.1e+03 & 7.56 & 40.5 & 1.22e-15 & -9.09 & 11.8 & 1.26e+05 & 9.68 & 7.21 & 6.89\\
32 & 6.21e+04 & 9.3 & 7.9 & 3.54e-12 & -7.12 & 58.1 & 5.23e-16 & -9.45 & 12.2 & 6e+03 & 7.32 & 42.5 & 7.07\\
33 & 1.86e-16 & -9.86 & 12.8 & 2.81e+04 & 8.91 & 8.68 & 3.98e+03 & 7.13 & 44.5 & 8.01e-12 & -6.89 & 58.7 & 7.28\\
34 & 9.95e+03 & 8.43 & 9.54 & 4.7e-17 & -10.4 & 13.5 & 2.66e+03 & 6.97 & 46.7 & 1.19e-11 & -6.79 & 61 & 7.55\\
35 & 2.92e+03 & 7.59 & 9.37 & 1.58e-10 & -6.07 & 60.4 & 4.05e-16 & -9.88 & 13.2 & 1.05e+03 & 6.33 & 46.2 & 7.21\\

>>>>

\usepackage{tikz}
\usepackage{bm}
\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}

@@

% $c_{d_{\text{h}}}$ \> horizontal tail profile drag coefficient \\

@@

% $c_{d_{\text{v}}}$ \> vertical tail profile drag coefficient \\

@@

$e$ \> span efficiency factor \\

@@

% $m_{\text{fac-wing}}$ \> margin factor\\
$m_{\text{fac}}$ \> weight margin factor \\

@@

$\mathcal{S}$ \> shear force \\ % [N] \\

@@

$W_{\text{fadd}}$ \> additional wing structural weight\\ % [N] \\

@@

This paper presents a physics-based optimization model that uses geometric programming as a systematic and rapid approach to evaluate trade offs between the two archictures for a given mission.  Geometric programming, a form of convex optimization, is chosen as a means of evaluating the design space because of its rapid solve time and guaranteed convergence to a global optimum.\cite{gp}
Models and equations representing the interaction of the various disciplines are expressed in a geometric programming form that are then combined to form an optimization model. 
The optimization is solved multiple times for different design parameters and requirements to observe changes in aircraft size and weight.

@@

Using this methodology, trade offs are observed between different aircraft configurations, power sources, and requirements.  
The results show that gas-powered aircraft can generally be built lighter and can fly faster and can therefore fly at higher latitudes and in higher percentile wind speeds.  
Solar-powered aircraft have greater endurance but are limited operationally by their ability to reach higher speeds.  
This optimization methodology can quantify the difference in weight and endurance between and gas and solar-electric powered aircraft for the same set of requirements. 
Varying mission requirements reveals how the best architecture changes depending on the requirements.

@@

Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
Therefore, the station keeping requirement is defined as the percentage of time the aircraft can remain on station, which is the percentile wind speed, $p_{\text{wind}}$. 
The wind speed at a station is also a function of latitude, altitude, season or day of the year,

@@

For long-endurance gas powered aircraft, the endurance requirement is defined by amount of time the aircraft needs to remain in the air without refueling.  
For a solar-electric powered aircraft, however, if it can fly during the shortest day of the year, December 21st (June 21st for Southern Hemisphere), then it can thoretically fly at any time of the year, more than satisfying the 6 day endurance requirement. 
Therefore, to meet the endurance requirement the solar-electric powered aircraft must be able to fly during the winter solstice.  
To achieve this design constraint the aircraft must be able to fly with limited solar energy, discussed in Section IV.A, and with higher than average wind speeds as shown in Figure~\ref{f:windvsmonth}.  
Because becuase wind speeds are greatest during the winter solstice for both the Northern and Southern Hemispheres only winds speeds from December and June respectively, will be used for the aircraft sizing and performance analysis presented in this paper.

\begin{figure}[H]
 \begin{subfigmatrix}{2}% number of columns
     \subfigure[-30$^{\circ}$ latitude \label{f:windvsmonthS30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthS30.pdf}}
     \subfigure[+30$^{\circ}$ latitude \label{f:windvsmonthN30}]{\includegraphics[natwidth=560,natheight=420]{windvsmonthN30.pdf}}
 \end{subfigmatrix}
 \caption{\textbf{Winds peak at the winter solstice in Southern and Northern Hemispheres.\cite{wind}}}
 \label{f:windvsmonth}
\end{figure}

@@

Long-endurance aircraft tend to fly at low velocities, which based on Figure~\ref{f:altvswindhmin} means there are two possible operating altitude regiems: at or around the minimum payload requirement 4,600 - 7,600 m or at high altitudes between 16,700-20,000 m. 
Gas powered aircraft generally fly in the lower altitude regime of 4,600-7,600 m because naturally aspirated engines lose power with increased altitude and are unable to reach altitudes higher than about 13,500 m.  
Solar-electric powered aircraft, which do not have naturally aspirated engines, will to fly around 18,000 m to avoid cloud coverage.

@@

Latitude additionally affects the solar-electric powered aircraft because at higher latitudes there is less daylight and therefore less solar energy during the winter months. This is further discussed in Section IV.A.

@@

Various structural, aerodynamic, and weight models necesistate assumptions about the vehicle configuration.  
Common characteristics between the two models include a constant tapered wing and a conventional tail with a single tail boom extending from the wing.  
The gas powered aircraft has a fuselage to hold all of the fuel required for the mission.  
The solar-electric powered aircraft holds the batteries in the wings.  
The solar cells for the solar-electric aircraft are placed along the wing and possibly on the horizontal tail.  
A simple diagram of what each vehicle might look like is shown in Figure~\ref{f:simpleaircraft} with their corresponding weight breakdowns in Figure~\ref{f:wbreak}.
The configuration is assumed to be static for this study.

@@

@@

Because wind speeds also vary with latitude, several constraints are needed to optimize altitude. 
Constraints relating wind speed and air density were generated using the data fitting technique described by Hoburg et. al\cite{fitting}.
Air density was chosen as the dependent variable instead of altitude to eliminate an extra equation relating altitude to air density. 
Altitude can be post-calculated from the optimized air density. (See Appendix A)

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the years 2005-2015.
In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix B) 
% The constraint on December wind speeds and air density at the 30 degree latitude is 
% 
% \begin{align}
%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
%     \label{e:windfit35}
%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
% \end{align}
% 
% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed. The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}. 
% 
% \begin{figure}[H]
% 	\begin{center}
% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
% 	\label{f:windfitl35}
% 	\end{center}
% \end{figure}

@@

\begin{align}
    \label{e:solarreq}
    (E/S)_{\text{sun}}  &\geq (E/S)_{\text{day}} + \frac{E_{\text{batt}}}{\eta_{\text{charge}}\eta_{\text{solar}} S_{\text{solar}}} \\
    \label{e:solarbatt}
    E_{\text{batt}} \eta_{\text{discharge}} &\geq P_{\text{oper}}t_{\text{night}} + (E/S)_{\text{twilight}} \eta_{\text{solar}} S_{\text{solar}}
\end{align}

@@

where $\eta_{\text{motor}} = 0.95$. 
Thus, for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude. 
$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix C to generate the fitted equations. 
% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
% 
% \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
%  \end{subfigmatrix}
%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
%  \label{f:energyapprox}
% \end{figure}

@@

\subsubsection{Motor Weight}

@@

This version comes from assuming that $\text{BSFC}$ and the power to weight ratio, $(P_{\text{shaft}}/W)$, are constant during the considered flight segment. 
One way to obtain a constant power to weight ratio is a constant velocity and constant lift coefficient.\cite{br2}
The average weight, $W_{\text{ave}}$, is assumed to be the geometric mean, defined as

@@

    To estimate the wing profile drag, a posynomial fit is made of the drag polar of the selected \emph{JH01} airfoil (See Appendix D for a discussion of the airfoil choice). 
    Drag polars were produced using XFOIL\cite{xfoil} at various Reynolds and the data was then fit to a posynomial equation using techniques described in Hoburg et. al.\cite{fitting}
    The XFOIL drag polar data is compared to the posynomial fit in Figure~\ref{f:JH01polar}.

%     \begin{align}
%         \label{e:aerodragprof}
%         c_{d_p}^{3.72} &\geq 0.0247C_L^{2.49}Re^{-1.11} + 2.03e^{-7}C_L^{12.7}Re^{-0.338} + 6.35e^{10}C_L^{-0.243}Re^{-3.43} \nonumber \\
%                        &+ 6.49e^{-6}C_L^{-1.9}Re^{-0.681} \\
%         Re &= \frac{\rho V S/b}{\mu}
%     \end{align}

@@

\item{Standard Wing Bending Case}

@@

\item{Gust Loading Case}

@@

The gust velocity is assumed to be vertical to the flight path such that the local angle of attack is approximated by

@@

\item{Discretized Beam Model}

@@

\item{Cap Spar for Bending Loads}

A cap spar is considered for the solar-electric and gas powered aircraft.  A cap spar has two carbon fiber caps separated by a foam core as seen in Figure~\ref{f:capspar}. A thin shear web is wrapped around the caps and foam to prevent shearing and buckling.

@@

\subsubsection{Additional Wing Weight}

@@

Additional wing weight, $W_{\text{fadd}}$, accounts for additional structural weight in the wing,

\begin{equation}
    W_{\text{fadd}} \geq (W_{\text{spar}} + W_{\text{skin}}) m_{\text{fac}}
\end{equation}

where $m_{\text{fac}} = 1.2$.

@@

The gas and solar-electric powered aircraft optimization models were solved by minimizing the take of weight for the requirements listed in Table~\ref{t:mreqs}. 
Key design parameters for both aircraft optimization model are listed in Table~\ref{t:gassolarparams}.
The gas powered aircraft optimization model had 552 unknowns and was solved in 0.144 seconds for a max take off weight of 84.0 lbs.  
The solar-electric powered optimization model had 375 unknowns and was solved in 0.164 seconds for a max take off weight of 320.5 lbs.
According to the assumptions described in the previous section, the gas powered aircraft is better suited to meet this specific set of requirements. 
This section explores how this result may vary for different requirements, input parameter values, and physics modeling assumptions.

\begin{longtable}{llll}
\caption{Fixed Optimization Parameters} \\
\toprule
\toprule
\multicolumn{2}{c}{Gas Powered} & \multicolumn{2}{c}{Solar Powered}\\
\midrule
$\eta_{\text{prop}}$         & 0.8  & $\eta_{\text{prop}}$         & 0.8 \\
$\tau$                       & 0.5  & $\tau$                       & 0.115 \\
$\lambda$                    & 0.5  & $\lambda$                    & 0.5 \\
$\sigma_{\text{CFRP}}$ [MPa] & 1700 & $\sigma_{\text{CFRP}}$ [MPa] & 1700  \\
$N_{\text{max}}$ - g-Loading & 5    & $N_{\text{max}}$ - g-Loading & 5 \\
$N_{\text{max}}$ - gust load & 2    & $N_{\text{max}}$ - gust load & 2 \\
$\text{BSFC}_{100\%}$ [kg/hr/kW] & 0.31 & $\eta_{\text{solar}}$      & 0.22 \\
Lapse Rate [hp/1000 ft]      & 0.035 &  $h_{\text{batt}}$ [Whr/kg]   & 350 \\
$\dot{h}$ [ft/min]           & 100 & $\eta_{\text{motor}}$           & 0.95 \\ 
$\rho_{\text{fuel}} $ [lbf/gallon] & 6.01 & $\eta_{\text{charge}}$          & 0.98 \\
                             &     & $\eta_{\text{discharge}}$       & 0.98 \\
\bottomrule
\label{t:gassolarparams}
 \end{longtable}

@@

Important trade offs between the gas powered and solar-electric powered archictures are highlighted and quantified by changing the latitude and endurance endurance requirements.   
Both optimization models were solved by minimizing the max take off weight across different bands of latitudes. 
Figure~\ref{f:latvsmtowtrade} shows this result evaluated at the 80th, 90th and 95th percentile wind speeds.  
The gas powered model was solved 63 times in 5.6178 seconds and the solar-electric powered model was solved 31 times in 2.5529 seconds to produce Figure~\ref{f:latvsmtowtrade}.

@@

While the solar-electric powered aircraft may be limited operationally by higher latitudes, it is not limited in endurance as is the gas powered aircraft.
Solving the gas powered optimization model for different endurance requirements shows where the gas powered archicture becomes less feasible. 
Figure~\ref{f:mtowvsendurance} shows the endurance versus size analysis for a gas powered aircraft by minimizing max take off weight for an aircraft capable of flying at any latitude. 
Figure~\ref{f:mtowvsendurance} was generated using 29 separate optimization solutions that took a total of 2.403 seconds to solve.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=527,natheight=405]{mtowvsendurance.pdf}
    \caption{\textbf{Endurance and size trade study for gas powered architecture.}}
	\label{f:mtowvsendurance}
	\end{center}
\end{figure}

@@

The previous results are dependent on the assumed input values and parameters.  
Changing the input values can help show where the designs becomes infeasible. 
Two input values that are especially important to the solar aircraft are the solar cell efficiency and battery specific energy. 
By solving the model for different assumed solar cell efficiency and battery specific energy values a broader picture of the design space is achieved.   
Figure~\ref{f:battsolarcontour} shows contours of latitude for a given solar cell efficiency and battery energy density.  
Put another way, this plot shows how good the solar cells and batteries must be in order to reach a given latitude. 
Figure~\ref{f:battsolarcontour} was produced using 157 separate optimiation solutions that took a total of 14.379 seconds to solve.

@@

By altering the aircraft structural model it can be observed how air density trades for wing weight. 
It might be assumed that because the wind speeds are lowest at 20,400 m at 29 degrees latitude, that the aircraft will always fly at 20,400 m.  
If it is assumed that the structural weight of the aircraft can be modeled as a fraction of the total weight

@@
